---
title: "7 - intercations"
author: "Aske Qvist"
date: "26 March 2020"
output: md_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 7 - interactions

Loading data
```{r}
# Packages
library(pacman)
p_load(rethinking, brms, tidyverse, patchwork)

# Data
data(rugged)
d <- rugged

detach(package:rethinking, unload = T)

```

Make data subsets
```{r}

# make log version of outcome
d <- d %>% mutate(
  log_gdp = log(rgdppc_2000)
)

# extract countries with GDP data
dd <- d %>% filter(complete.cases(rgdppc_2000))

# split countries into Africa and not-Africa
d.A1 <- dd %>% filter(cont_africa == 1) # Afrika
d.A0 <- dd %>% filter(cont_africa == 0) # not Afrika

```

Fit models to the two data subsets
```{r}

b7.1 <-
  brm(data = d.A1, family = gaussian, # Afrikan data
      log_gdp ~ 1 + rugged,
      prior = c(prior(normal(8, 100), class = Intercept),
                prior(normal(0, 1), class = b),
                prior(uniform(0, 10), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 7)

b7.2 <-
  update(b7.1, # Not Afrikan data
         newdata = d.A0)

```

Plot posterior predictions
```{r}
africa <- d.A1 %>%
  ggplot(aes(x = rugged, y = log_gdp)) +
  stat_smooth(method = "lm", fullrange = T, size = 1/2,
              color = "firebrick4", fill = "firebrick", alpha = 1/5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 1/2) +
  scale_x_continuous("Rugged", limits = c(0, 55)) +
  coord_cartesian(xlim = 0:7, ylim = 6:10) +
  ylab("Log of GDP") +
  theme_bw() + ggtitle("Africa") 
  
not_africa <- d.A0 %>%
  ggplot(aes(x = rugged, y = log_gdp)) +
  stat_smooth(method = "lm", fullrange = T, size = 1/2,
              color = "firebrick4", fill = "firebrick", alpha = 1/5) +
  geom_point(size = 1.5, color = "firebrick4", alpha = 1/2) +
  scale_x_continuous("Rugged", limits = c(0, 55)) +
  coord_cartesian(xlim = 0:7, ylim = 6:10) +
  ylab("Log of GDP") +
  theme_bw() + ggtitle("Not Africa")
  

not_africa + africa
# slope is positive within africa but negative outside africa


```

It is better if we keep the data in the same data set. Multilevel models borrow data from the different categories in order to improve estimates in all categories - and many other reasons. But even if we added the dummy variable cont_africa, this will not solve the problem - as a normal linear regression does not account for different slopes for the categories of the dummy.

## Comapre the models with/without the dummy cont_africa as predictor ( but no interaction effect)
```{r}
# Start with the model I made before (without the dummy cont_africa) - just with all the data
b7.3 <-
  update(b7.1, newdata = dd) # all data

# Now the same but added is the dummy as a predictor (in the formula)
b7.4 <-
  update(b7.3,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged + cont_africa) 

# compute information criteria for both
b7.3 <- add_criterion(b7.3, c("loo", "waic"))
b7.4 <- add_criterion(b7.4, c("loo", "waic"))


# compare IC
loo_compare(b7.3, b7.4,
            criterion = "waic")

loo_compare(b7.3, b7.4,
            criterion = "loo")
# loo and waic agree - the model with
# the dummy fits teh data way better


# compute weight
model_weights(b7.3, b7.4,
              weights = "waic") %>% 
  round(digits = 3)
# The model with the dummy 
# has all the weight :)


```


## adding a linear interaction
We want the realtionship between Y (GDP) and R (rugged) to vary as a function of A (dummy varible)

(the model comparisons do not work) add_criterion() and model_weigths() when using brms.
```{r}

# Define new model with interaction effect 
b7.5 <-
  update(b7.4,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged*cont_africa) # syntax as lmer()

b7.5b <-
  update(b7.4,
         newdata = dd,
         formula = log_gdp ~ 1 + rugged + cont_africa + rugged:cont_africa) 

# add IC 
b7.5 <- add_criterion(b7.5,c("loo","waic"))

# compare all 3 models
l <- loo_compare(b7.3, b7.4, b7.5, criterion = "waic")

print(l, simplify = F)

# weight
model_weights(b7.3, b7.4, b7.5,
              weights = "loo") %>% 
  round(digits = 3)

b7.5$loo # This is not working!



```


## Interpreting interaction effect 
and compare and plot slopes for the dummy
```{r}
## Parameter estimates 
posterior_summary(b7.5)
posterior_summary(b7.4)
# Does not calculate the gamma (the slope conditioned on cont_africa). Therefore, we have to do it manually

# within Africa
fixef(b7.5)[2, 1] + fixef(b7.5)[4, 1] * 1
# outside Africa
fixef(b7.5)[2, 1] + fixef(b7.5)[4, 1] * 0

#But those are only map values, from a sample of the posterior distribution we can get uncertainties around those MAP's
post <- posterior_samples(b7.5) 

post %>%
  transmute(gamma_Africa    = b_rugged + `b_rugged:cont_africa`, # based on computing gamma when A == 1
            gamma_notAfrica = b_rugged) %>% # based on computing gamma when A == 0 (and the interaction effect disappears)
  gather(key, value) %>%
  group_by(key) %>%
  summarise(mean = mean(value))


# plot the distributions (of uncertainty around the MAP values - the slopes for the dummy)
p_load(ggthemes)

post %>%
  transmute(gamma_Africa    = b_rugged + `b_rugged:cont_africa`,
            gamma_notAfrica = b_rugged) %>%
  gather(key, value) %>%
  
  ggplot(aes(x = value, group = key, color = key, fill = key)) +
  geom_density(alpha = 1/4) +
  scale_colour_pander() +
  scale_fill_pander() +
  scale_x_continuous(expression(gamma), expand = c(0, 0)) +
  scale_y_continuous(NULL, breaks = NULL) +
  ggtitle("Terraine Ruggedness slopes",
          subtitle = "Blue = African nations, Green = others") +
  theme_pander() + 
  theme(text = element_text(family = "Times"),
        legend.position = "none")
# nations outside Africa has a positive relationship between GDP and rugged, with a distribution of uncertainty around

#what’s the probability (according to this model and these data) that
#the slope within Africa is less than the slope outside of Africa?

post %>%
  mutate(gamma_Africa    = b_rugged + `b_rugged:cont_africa`,
         gamma_notAfrica = b_rugged) %>% 
  mutate(diff            = gamma_Africa - gamma_notAfrica) %>% 
  # what proportion of the differences (for each sample of the posterior) is below 0?
  summarise(Proportion_of_the_difference_below_0 = sum(diff < 0) / length(diff))

```



## CONTINUOUS INTERACTIONS

load data
```{r}
# The data
library(rethinking)
data(tulips)
d <- tulips
head(d)

# done in online version
detach(package:rethinking, unload = T)
library(brms)
rm(tulips)

```

## making models with interaction to compare

Making 2 models - one where the parameters interact and one without interaction
```{r}
# Without
b7.6 <-
  brm(data = d, family = gaussian,
      blooms ~ 1 + water + shade,
      prior = c(prior(normal(0, 100), class = Intercept),
                prior(normal(0, 100), class = b),
                prior(uniform(0, 100), class = sigma)),
      iter = 2000, warmup = 1000, cores = 4, chains = 4,
      seed = 7)

# yieled divergent transitions --> update with better prior and 'control' element (whatever that means)

b7.6 <-
  update(b7.6,
         prior = c(prior(normal(0, 100), class = Intercept),
                   prior(normal(0, 100), class = b),
                   prior(cauchy(0, 10), class = sigma)),
         control = list(adapt_delta = 0.9), # apparently good to add.
         seed = 7)

posterior_summary(b7.6) %>% round(digits = 2)


# with interaction effect
b7.7 <- 
  update(b7.6, 
         formula = blooms ~ 1 + water + shade + water:shade)

posterior_summary(b7.6) %>% round(digits = 2)
posterior_summary(b7.7) %>% round(digits = 2)


# model comparison with WAIC
b7.6 <- add_criterion(b7.6, "waic")
b7.7 <- add_criterion(b7.7, "waic")

w <- loo_compare(b7.6, b7.7, criterion = "waic")

print(w, simplify = F)

# cbind() trick to convert from elpd metric to more traditional WAIC metric
cbind(waic_diff = w[, 1] * -2,
      se        = w[, 2] *  2)

# and weight 
model_weights(b7.6, b7.7, weights = "waic") # almost all weight goes to interaction model.

```

I do it after mean centering
```{r}
# mean center to two new columns
# it is better to meancenter beacuse then the intercept will be the empirical mean - and the search when compiling the model will work better(i think)
d <- d %>% mutate(
  water_c = water - mean(water),
  shade_c = shade - mean(shade)
)

# compile new model without interaction
b7.8 <-
  brm(data = d, family = gaussian,
      blooms ~ 1 + water_c + shade_c,
      prior = c(prior(normal(130, 100), class = Intercept),# updating prior for intercept (after summary, I know the mean
                prior(normal(0, 100), class = b),          # of outcome varible is around 130)
                prior(cauchy(0, 10), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      control = list(adapt_delta = 0.9),
      seed = 7)

# with interaction
b7.9 <- 
  update(b7.8, 
         formula = blooms ~ 1 + water_c + shade_c + water_c:shade_c)


# Results
posterior_summary(b7.8) %>% round(digits = 2)
posterior_summary(b7.9) %>% round(digits = 2)
# because we have mean centered the predictor variables,
# the intercept becomes the grand mean of the outcome



# Model comparison 
b7.8 <- add_criterion(b7.8, "waic")
b7.9 <- add_criterion(b7.9, "waic")

h <- loo_compare(b7.8, b7.9, criterion = "waic")

print(h, simplify = F)

# cbind() trick to convert from elpd metric to more traditional WAIC metric
cbind(waic_diff = w[, 1] * -2,
      se        = w[, 2] *  2)

# and weight 
model_weights(b7.8, b7.9, weights = "waic") # most weight ascribed to b7.9


# comparative summary of the 2 models
tibble(model  = str_c("b7.", 8:9)) %>% 
  mutate(fit  = purrr::map(model, get)) %>% 
  mutate(tidy = purrr::map(fit, broom::tidy)) %>% 
  unnest(tidy) %>% 
  filter(term != "lp__") %>% 
  select(term, estimate, model) %>% 
  spread(key = model, value = estimate) %>% 
  mutate_if(is.double, round, digits = 2)


```

Plot implied predictions - TRIPTYCH (good way to plot interactions)
```{r}
# three plot made to be viewed together


# loop over values of water_c and plot predicitons
shade_seq <- -1:1

for(w in -1:1) {
  # define the subset of the original data
  dt <- d[d$water_c == w, ]
  # defining our new data
  nd <- tibble(water_c = w, shade_c = shade_seq)
  # use our sampling skills, like before
  f <- 
    fitted(b7.9, newdata = nd) %>%
    as_tibble() %>%
    bind_cols(nd)
  
  # specify our custom plot
  fig <- 
    ggplot() +
    geom_smooth(data = f,
                aes(x = shade_c, y = Estimate, ymin = Q2.5, ymax = Q97.5),
                stat = "identity", 
                fill = "#CC79A7", color = "#CC79A7", alpha = 1/5, size = 1/2) +
    geom_point(data = dt, 
               aes(x = shade_c, y = blooms),
               shape = 1, color = "#CC79A7") +
    coord_cartesian(xlim = range(d$shade_c), 
                    ylim = range(d$blooms)) +
    scale_x_continuous("Shade (centered)", breaks = c(-1, 0, 1)) +
    labs("Blooms", 
         title = paste("Water (centered) =", w)) +
    theme_pander() + 
    theme(text = element_text(family = "Times"))

  # save to plotting
  if (w == -1){
    fig_1 <- fig
  } else if (w == 0){
    fig_2 <- fig
  } else {
    fig_3 <- fig
  }

}

# Using patchwork to plot them next to eachother
fig_1 + fig_2 + fig_3

```








# Exercises for chapter 7


## Easy

### 7E1. For each of the causal relationships below, name a hypothetical third variable that would lead to an interaction effect.

(1) Bread dough rises because of yeast.
- the continues variable of temperature interacting with the amount of yeast

(2) Education leads to higher income.
- Gender, colour etc.

(3) Gasoline makes a car go.
- number of cylinders

### 7E2. Which of the following explanations invokes an interaction?

(1) Caramelizing onions requires cooking over low heat and making sure the onions do not
dry out.
- Unclear 

(2) A car will go faster when it has more cylinders or when it has a better fuel injector.
- Interaction

(3) Most people acquire their political beliefs from their parents, unless they get them instead
from their friends.
- No interaction


(4) Intelligent animal species tend to be either highly social or have manipulative appendages
(hands, tentacles, etc.).
- No interaction

### 7E3. For each of the explanations in 7E2, write a linear model that expresses the stated relationship.
```{r}
onion ~ 1 + heat * awareness
onion ~ heat + dryness

speed ~ 0 + num_cylinders:fuel_injector + num_cylinders + fuel_injector
# Same as: 
speed ~ 1 + num_cylinders * fuel_injector

pol_belief ~ parents + friends

intelligenec ~ social + mani_appendage


```

## Medium.
7M1. Recall the tulips example from the chapter. Suppose another set of treatments adjusted the
temperature in the greenhouse over two levels: cold and hot. The data in the chapter were collected
at the cold temperature. You find none of the plants grown under the hot temperature developed
any blooms at all, regardless of the water and shade levels. Can you explain this result in terms of
interactions between water, shade, and temperature?
```{r}
BloomSize ~ (Water * Shade) * Temperature

```

7M2. Can you invent a regression equation that would make the bloom size zero, whenever the
temperature is hot?

Temperature is a binary variable, where 0 is hot and 1 is colder.

$$
BloomSize \sim  (a + bW*water.c + bS*shade.c + bWS*water.c*shade.c)*Temp
\\
or
\\
BloomSize = (a + bW*water.c + bS*shade.c + bWS*water.c*shade.c)*Temp + \epsilon
$$

7M3. In parts of North America, ravens depend upon wolves for their food. This is because ravens
are carnivorous but cannot usually kill or open carcasses of prey. Wolves however can and do kill
and tear open animals, and they tolerate ravens co-feeding at their kills. This species relationship
is generally described as a "species interaction". 

a)
Can you invent a hypothetical set of data on raven
population size in which this relationship would manifest as a statistical interaction? 
```{r}
pacman::p_load(tidyverse, patchwork, lme4)

food <- c(rep(1,5),rep(2,5),rep(3,5))
wolves <- c(12,13,11,10,9,24,22,25,27,16,27,29,30,28,30)
ravens <- c(0,1,0,2,3,5,6,7,4,5,25,27,31,30,29)

d <- as.data.frame(cbind(food,wolves,ravens))

food1 <- ggplot(data = d[d$food == 1,], aes(x = wolves, y = ravens)) + geom_smooth(method = 'lm') + geom_point() + labs(title = 'Food low')
food2 <- ggplot(data = d[d$food == 2,], aes(x = wolves, y = ravens)) + geom_smooth(method = 'lm') + geom_point() + labs(title = 'Food relatively medium')
food3 <- ggplot(data = d[d$food == 3,], aes(x = wolves, y = ravens)) + geom_smooth(method = 'lm') + geom_point() + labs(title = 'Food high')

food1 + food2 + food3

RavenPopuSize ~ PreyPopulationWolvesPopulation

raven ~ normal(mu,sigma)
mu ~ food:wolves + wolves + food

summary(lm(ravens ~ wolves * food), data = d)
summary(lm(ravens ~ 0 + wolves * as.factor(food)), data = d)


```
b)
Do you think the biological interaction could be linear? Why or why not?

```{r}
# It might be.
# But it might also be the case that if the ratio between Prey and Wolves increase, the increase of the interaction effect on Ravens increase. 

# I.e. That the more prey there is per wolf, the less of the meat they eat from the prey -> thus leaving more prey for the ravens (leading the population of ravens to increase).


```


## Hard.
7H1. Return to the data(tulips) example in the chapter. Now include the bed variable as a predictor
in the interaction model. Don't interact bed with the other predictors; just include it as a main
effect. Note that bed is categorical. So to use it properly, you will need to either construct dummy
variables or rather an index variable, as explained in Chapter 6.
```{r}

# The data
library(pacman)
p_load(brms,tidyverse)
library(rethinking)
data(tulips)
d <- tulips
head(d)

# done in online version
detach(package:rethinking, unload = T)

# construct dummy variable
d$bed <- as.factor(d$bed)



# mean centering
d <- d %>% mutate(
  water_c = water - mean(water),
  shade_c = shade - mean(shade)
)


# without interaction with bed
#formula
f_7h1 <- bf(blooms ~ 1 + water_c * shade_c + bed)

#getting prior
get_prior(f_7h1, data=d, family=gaussian)

#defining priors
prior_f_7h1 <- c(prior(normal(0, 100), class = Intercept),
                prior(normal(0, 100), class = b),
                prior(uniform(0, 100), class = sigma)) 

#model
m_7h1 <- 
  brm(data = d, family = gaussian,
     f_7h1,
      prior = prior_f_7h1,
      iter = 2000, 
     warmup = 1000, 
     cores = 2, 
     chains = 2,
      seed = 7,
     sample_prior = T)

#posterior check
pp_check(m_7h1, nsamples = 100)

#summary
m_7h1

# add criterion
m_7h1 <- add_criterion(m_7h1, "loo")

```

7H2. Use WAIC to compare the model from 7H1 to a model that omits bed. What do you infer
from this comparison? Can you reconcile the WAIC results with the posterior distribution of the bed
coefficients?
```{r}
# with interaction with bed
#formula
f_7h2 <- bf(blooms ~ 1 + water_c * shade_c)

#getting prior
get_prior(f_7h2, data=d, family=gaussian)

#defining priors
prior_f_7h2 <- c(prior(normal(0, 100), class = Intercept),
                prior(normal(0, 100), class = b),
                prior(uniform(0, 100), class = sigma)) 

#model
m_7h2 <- 
  brm(data = d, family = gaussian,
     f_7h2,
      prior = prior_f_7h2,
      iter = 2000, 
     warmup = 1000, 
     cores = 2, 
     chains = 2,
      seed = 7,
     sample_prior = T)

#posterior check?
pp_check(m_7h2, nsamples = 100)

# add criterion
m_7h2 <- add_criterion(m_7h2, "loo", reloo = TRUE)

#comparing the two with LOOIC weights

loo_compare(m_7h2, m_7h1)
loo_model_weights(m_7h2, m_7h1)





# cbind() trick to convert from elpd metric to more traditional WAIC metric
cbind(waic_diff = w[, 1] * -2,
      se        = w[, 2] *  2)

# and weight 
model_weights(b7.6, b7.7, weights = "waic") # almost all weight goes to interaction model.


```

7H3. Consider again the data(rugged) data on economic development and terrain ruggedness,
examined in this chapter. One of the African countries in that example, Seychelles, is far outside
the cloud of other nations, being a rare country with both relatively high GDP and high ruggedness.
Seychelles is also unusual, in that it is a group of islands far from the coast of mainland Africa, and
its main economic activity is tourism.
One might suspect that this one nation is exerting a strong influence on the conclusions. In
this problem, I want you to drop Seychelles from the data and re-evaluate the hypothesis that the
relationship of African economies with ruggedness is different from that on other continents.

(a) Begin by using map to fit just the interaction model:
$$y_i ~ Normal(\mu_i, \sigma)$$
$$\mu_i = \alpha + \beta A_i + \beta_R R_i + \beta_AR  A_i R_i$$
where y is log GDP per capita in the year 2000 (log of rgdppc_2000); A is cont_africa, the dummy
variable for being an African nation; and R is the variable rugged. Choose your own priors. Compare 
the inference from this model fit to the data without Seychelles to the same model fit to the full data.
Does it still seem like the effect of ruggedness depends upon continent? How much has the expected
relationship changed?

```{r}

```

(b) Now plot the predictions of the interaction model, with and without Seychelles. Does it still
seem like the effect of ruggedness depends upon continent? How much has the expected relationship
changed?

```{r}

```

(c) Finally, conduct a model comparison analysis, using WAIC. Fit three models to the data without
Seychelles:

$$Model 1 : y_i ~ Normal(\mu_i, \sigma) 
\mu_i = \alpha + \beta_RR_i $$

$$Model 2 : y_i ~ Normal(\mu_i, \sigma)      

\mu_i = \beta_A A_i + \beta_R R_i$$

$$Model 3 : y_i ~ Normal(\mu_i, \sigma)
\mu_i = \beta_A A_i + \beta_R R_i + \beta_AR  A_i R_i$$


```{r}

```










