---
title: "Class 7 - interaction exercises"
author: "Rebecca Foss"
date: "4/2/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse, brms, rethinking)

```

# Easy
7E1. For each of the causal relationships below, name a hypothetical third variable that would lead to an interaction effect.
(1) Bread dough rises because of yeast.
Temperature 
(2) Education leads to higher income.
Job, gender, race
(3) Gasoline makes a car go.
Number of cylinders, slope of the road


7E2. Which of the following explanations invokes an interaction?
(1) Caramelizing onions requires cooking over low heat and making sure the onions do not
dry out.
Up for interpretation 

(2) A car will go faster when it has more cylinders or when it has a better fuel injector.
Has to interact. Speed of car is dependant on number of cylinders and fuel injector

(3) Most people acquire their political beliefs from their parents, unless they get them instead
from their friends.
No interaction

(4) Intelligent animal species tend to be either highly social or have manipulative appendages
(hands, tentacles, etc.).
No interaction

7E3. For each of the explanations in 7E2, write a linear model that expresses the stated relationship.
```{r}
CaramelizedOnions ~ 
```

# Medium.
7M1. Recall the tulips example from the chapter. Suppose another set of treatments adjusted the
temperature in the greenhouse over two levels: cold and hot. The data in the chapter were collected
at the cold temperature. You find none of the plants grown under the hot temperature developed
any blooms at all, regardless of the water and shade levels. Can you explain this result in terms of
interactions between water, shade, and temperature?
```{r}

```

7M2. Can you invent a regression equation that would make the bloom size zero, whenever the
temperature is hot?
```{r}

```

7M3. In parts of North America, ravens depend upon wolves for their food. This is because ravens
are carnivorous but cannot usually kill or open carcasses of prey. Wolves however can and do kill
and tear open animals, and they tolerate ravens co-feeding at their kills. This species relationship
is generally described as a “species interaction.” Can you invent a hypothetical set of data on raven
population size in which this relationship would manifest as a statistical interaction? Do you think
the biological interaction could be linear? Why or why not?
```{r}

```

# Hard.
7H1. Return to the data(tulips) example in the chapter. Now include the bed variable as a predictor
in the interaction model. Don’t interact bed with the other predictors; just include it as a main
effect. Note that bed is categorical. So to use it properly, you will need to either construct dummy
variables or rather an index variable, as explained in Chapter 6.
```{r}

```

7H2. Use WAIC to compare the model from 7H1 to a model that omits bed. What do you infer
from this comparison? Can you reconcile the WAIC results with the posterior distribution of the bed
coefficients?
```{r}

```

7H3. Consider again the data(rugged) data on economic development and terrain ruggedness,
examined in this chapter. One of the African countries in that example, Seychelles, is far outside
the cloud of other nations, being a rare country with both relatively high GDP and high ruggedness.
Seychelles is also unusual, in that it is a group of islands far from the coast of mainland Africa, and
its main economic activity is tourism.
One might suspect that this one nation is exerting a strong influence on the conclusions. In
this problem, I want you to drop Seychelles from the data and re-evaluate the hypothesis that the
relationship of African economies with ruggedness is different from that on other continents.

(a) Begin by using map to fit just the interaction model:
$$y_i ~ Normal(\mu_i, \sigma)$$
$$\mu_i = \alpha + \beta A_i + \beta_R R_i + \beta_AR  A_i R_i$$
where y is log GDP per capita in the year 2000 (log of rgdppc_2000); A is cont_africa, the dummy
variable for being an African nation; and R is the variable rugged. Choose your own priors. Compare 
the inference from this model fit to the data without Seychelles to the same model fit to the full data.
Does it still seem like the effect of ruggedness depends upon continent? How much has the expected
relationship changed?

```{r}

```

(b) Now plot the predictions of the interaction model, with and without Seychelles. Does it still
seem like the effect of ruggedness depends upon continent? How much has the expected relationship
changed?

```{r}

```

(c) Finally, conduct a model comparison analysis, using WAIC. Fit three models to the data without
Seychelles:

$$Model 1 : y_i ~ Normal(\mu_i, \sigma) 
\mu_i = \alpha + \beta_RR_i $$

$$Model 2 : y_i ~ Normal(\mu_i, \sigma)      

\mu_i = \beta_A A_i + \beta_R R_i$$

$$Model 3 : y_i ~ Normal(\mu_i, \sigma)
\mu_i = \beta_A A_i + \beta_R R_i + \beta_AR  A_i R_i$$


```{r}

```



