---
title: "Computational Modeling - Assignment 2"
author: "Riccardo Fusaroli"
date: "29/01/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pacman)
p_load(tidyverse, ggplot2,rethinking,brms)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. there is a second part at the bottom for next week.

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

```{r}
data <- data.frame(
  correct=c(3,2,160,66),
  questions=c(6,2,198,132),
  teacher=c("RF","KT","JS","MW"))

```

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results

```{r}
#Grid Approximation
dens <- 20
p_grid <- seq(from = 0, to = 1, length.out = dens)
prior <- rep(1,dens)
#checking the prior in a density plot
dens(rbinom(1e4, 6, runif(1e4, 0, 1))) #Flat prior = the peaaks are all equally valid. y-axis = likelihood x = scale of the outcome
dens(rbinom(1e4, 6, runif(1e4, 0.5, 1))) #UnreasonablyOptimisticPrior = only above chance are waeighted more likely
dens(rbinom(1e4, 6, rnorm(1e4, 0.5, 0.1)))
#calculating likelihood
(likelihood <-dbinom( 3 , size = 6 , prob = p_grid)) #given the data what is the probability
plot(likelihood)
#calculate posterior
(unstd.posterior<-likelihood * prior)
plot(unstd.posterior)
#standadize posterior
posterior <-unstd.posterior/ sum(unstd.posterior)
#making a df of all the parameters
#from slides
d <-data.frame(grid = p_grid, posterior = posterior, prior = prior, likelihood = likelihood)

# #from brms
# grid_f <- function(data,dens){
#   for(i in 1:nrow(data))
#     {
#     correct <- data[i,1]
#     questions <- data[i,2]
#     teacher <- data[i,3]
#     
#     d2 <-
#     tibble(p_grid            = seq(from = 0, to = 1, length.out = dens),  # define grid
#            prior             = rep(1,dens)) %>%                         # define prior
#     mutate(likelihood      = dbinom(correct, size = questions, prob = p_grid)) %>%  # compute likelihood at each value in grid
#     mutate(unstd_posterior = likelihood * prior) %>%                  # compute product of likelihood and prior
#     mutate(posterior       = unstd_posterior / sum(unstd_posterior)) %>%   # standardize the posterior, so it sums to 1
#     mutate(peak_performance = match(max(posterior),posterior)/length(posterior)) %>% 
#     mutate(above_chance = sum( posterior[ p_grid > 0.5 ] ))
# 
# ggplot(d2, aes(grid,posterior)) +
#           geom_point() +geom_line()+theme_classic() +
#           geom_line(aes(grid, prior/dens),color= 'red')+
#           ggtitle(teacher) +
#           xlab("Knowledge of CogSci")+ ylab("posterior probability")
# }}

grid_f(data,100)
#plotting the grid approximation
ggplot(d2, aes(grid,posterior)) +
          geom_point() +geom_line()+theme_classic() +
          geom_line(aes(grid, prior/dens),color= 'red')+
          ggtitle(teacher) +
          xlab("Knowledge of CogSci")+ ylab("posterior probability")

#summarise 
match(max(d2$posterior),d2$posterior)/length(d2$posterior)
#peak probability is 0.5

# add up posterior probability where p > 0.5 
sum( d2$posterior[ d2$p_grid > 0.5 ] )
#the probability is 0.5 that he will answer more that 50% correct

```
Riccardo has 50% knowledge of cogsci
There is a 50% probability that Riccardo will answer above chance

- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)
```{r}
#Quadradic approximation
rf_qa <-
  rethinking::map(
    alist(
      a ~ dbinom(6, p),  # binomial likelihood
      p ~ dunif(0, 1)    # uniform prior
    ), 
    data = list(a = 3))

#summary of qa
precis(rf_qa)

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.
```{r}

g_approx_uniform <- function(data, dens){
  #Loop through data frame
  for(i in 1:nrow(data))
    {
    correct <- data[i,1]
    questions <- data[i,2]
    teacher <- data[i,3]
  #Define grid
  prob_grid <- seq(from = 0 , to = 1 , length.out = dens)
  #Prior
  prior <- rep(1, dens)
  #Likelihood
  likelihood <- dbinom(correct, size = questions, prob = prob_grid)
  #Posterior
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  #Plot
  d <- data.frame(grid = prob_grid, posterior = posterior, prior = prior, likelihood = likelihood)
  print(ggplot(d, aes(grid,posterior)) +
          geom_point() +geom_line()+theme_classic() +
          geom_line(aes(grid, prior/dens),color= 'red')+
          ggtitle(teacher) +
          xlab("Knowledge of CogSci")+ ylab("posterior probability"))
  #Summary outcome
  teacher <- as.character(teacher)
  s <- c((match(max(posterior),posterior)/length(posterior)),sum(posterior[prob_grid > 0.5 ] ),teacher)
  print(s)
}}

g_approx(data,100)

calc_teacher <- function(teacher, correct, questions, prior, prob_grid){
   
   # Compute likelihood 
   likelihood <- dbinom( correct, size = questions, prob = prob_grid )
   
   # Compute unstandardized posterior from likelihood and the prior
   uns_posterior <- likelihood * prior
   
   # Compute standardized posterior. Now, it takes the bin size into account.
   posterior <- uns_posterior / sum(uns_posterior)
   
   # Compute MAP (Maximum a posteriori)
   map <- match(max(posterior),posterior) / length(posterior)
   
   # posterior probability where p > 0.5
   chance <- sum(posterior[ prob_grid > 0.5 ])
   
   # Teacher as factor
   teacher <- as.factor(teacher)
  
   # specify output
   return(list(teacher, map, chance, prior, likelihood, posterior))
   
 }

#Loop through all teachers
# Making empty dataframe
teacher_info <- data.frame(teacher = factor(), MAP = numeric(), chance = numeric(), prior = numeric(), likelihood = numeric(), posterior = numeric())
# Run loop to extract MAP and 'chance above chance' for each teacher
for(i in 1:nrow(data)) {
    
    correct <- data[i,1]
    questions <- data[i,2]
    teacher <- data[i,3]
    # Define grid and prior
    prob_grid <- seq(from = 0, to = 1, length.out = 10000)
    prior <- rep(1, 1e4)
    
    
    # Use my sexy little function
    info <- calc_teacher(teacher, correct, questions, prior, prob_grid) %>% as.data.frame()
    
    names(info)[1] <- "teacher"
    names(info)[2] <- "MAP"
    names(info)[3] <- "chance"
    names(info)[4] <- "prior"
    names(info)[5] <- "likelihood"
    names(info)[6] <- "posterior"
    
    # Combine with premade empty dataframe
    if (nrow(teacher_info) == 0) {
      teacher_info <- info}
      else {
        teacher_info <- rbind(teacher_info, info)}
    
    }

p_grid <- seq(from = 0, to = 1, length.out = 10000)


post_RF <- teacher_info$posterior[teacher_info$teacher=="RF"]
post_KT <- teacher_info$posterior[teacher_info$teacher=="KT"]
post_JS <- teacher_info$posterior[teacher_info$teacher=="JS"]
post_MW <- teacher_info$posterior[teacher_info$teacher=="MW"]

samples_RF <- sample(p_grid, prob=post_RF , size=1e4, replace = TRUE )
samples_KT <- sample(p_grid, prob=post_KT , size=1e4, replace = TRUE )
samples_JS <- sample(p_grid, prob=post_JS , size=1e4, replace = TRUE )
samples_MW <- sample(p_grid, prob=post_MW , size=1e4, replace = TRUE )

teacher_info$samples[teacher_info$teacher=="RF"] <- samples_RF
teacher_info$samples[teacher_info$teacher=="KT"] <- samples_KT
teacher_info$samples[teacher_info$teacher=="JS"] <- samples_JS
teacher_info$samples[teacher_info$teacher=="MW"] <- samples_MW

#Josh knowledge 
sum(samples_JS>samples_RF)/length(samples_JS)*100
sum(samples_JS>samples_KT)/length(samples_JS)*100
sum(samples_JS>samples_MW)/length(samples_JS)*100

#Riccardos knowledge
sum(samples_RF>samples_JS)/length(samples_RF)*100
sum(samples_RF>samples_KT)/length(samples_RF)*100
sum(samples_RF>samples_MW)/length(samples_RF)*100

#Kristians knowledge
sum(samples_KT>samples_RF)/length(samples_KT)*100
sum(samples_KT>samples_JS)/length(samples_KT)*100
sum(samples_KT>samples_MW)/length(samples_KT)*100

#Mikkels knowledge
sum(samples_MW>samples_JS)/length(samples_MW)*100
sum(samples_MW>samples_RF)/length(samples_MW)*100
sum(samples_MW>samples_KT)/length(samples_MW)*100
```
 

3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.
```{r}
##Changing the prior
g_approx_norm <- function(data, dens){
  #Loop through data frame
  for(i in 1:nrow(data))
    {
    correct <- data[i,1]
    questions <- data[i,2]
    teacher <- data[i,3]
  #Define grid
  prob_grid <- seq(from = 0 , to = 1 , length.out = dens)
  #Prior
  prior <- dnorm(prob_grid, 0.8, 0.2) #Normal with a SD of 0.2
  #Likelihood
  likelihood <- dbinom(correct, size = questions, prob = prob_grid)
  #Posterior
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  #Plot
  d <- data.frame(grid = prob_grid, posterior = posterior, prior = prior, likelihood = likelihood)
  print(ggplot(d, aes(grid,posterior)) +
          geom_point() +geom_line()+theme_classic() +
          geom_line(aes(grid, prior/dens),color= 'red')+
          ggtitle(teacher) +
          xlab("Knowledge of CogSci")+ ylab("posterior probability"))
  #Summary outcome
  teacher <- as.character(teacher)
  s <- c((match(max(posterior),posterior)/length(posterior)),sum(posterior[prob_grid > 0.5 ] ),teacher)
  print(s)
}}

g_approx_norm(data,100)

```

4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?
```{r}
dataBIG <- data %>% 
  mutate(correct = correct*100,
         questions = questions*100)

#For uniform distribution
g_approx_uniform(dataBIG, 1000)
#For normal distribution
g_approx_norm(dataBIG, 1000)

```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?
```{r}
#Conservative prior

g_approx_cons <- function(data, dens){
  #Loop through data frame
  for(i in 1:nrow(data))
    {
    correct <- data[i,1]
    questions <- data[i,2]
    teacher <- data[i,3]
  #Define grid
  prob_grid <- seq(from = 0 , to = 1 , length.out = dens)
  #Prior
  # prior <- exp(-abs(probability - .5) / .25) / ( 2 * .25) #Laplace distribution
  prior <- dnorm(prob_grid, 0.5, 0.01) #Normal with a SD of 0.05
  #Likelihood
  likelihood <- dbinom(correct, size = questions, prob = prob_grid)
  #Posterior
  unstd.posterior <- likelihood * prior
  posterior <- unstd.posterior / sum(unstd.posterior)
  #Plot
  d <- data.frame(grid = prob_grid, posterior = posterior, prior = prior, likelihood = likelihood)
  print(ggplot(d, aes(grid,posterior)) +
          geom_point() +geom_line()+theme_classic() +
          geom_line(aes(grid, prior/dens),color= 'red')+
          ggtitle(teacher) +
          xlab("Knowledge of CogSci")+ ylab("posterior probability"))
  #Summary outcome
  teacher <- as.character(teacher)
  s <- c((match(max(posterior),posterior)/length(posterior)),sum(posterior[prob_grid > 0.5 ] ),teacher)
  print(s)
  }}

#loop with new prior
g_approx_cons(dataBIG,1000)
```

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)

d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

# Model sampling only from the prior (for checking the predictions your prior leads to)
FlatModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = "only") # here we tell the model to ignore the data

# Plotting the predictions of the model (prior only) against the actual data
pp_check(FlatModel_priorCheck, nsamples = 100)

# Model sampling by combining prior and likelihood
FlatModel <- brm(Correct|trials(Questions) ~ 1, 
                 data = subset(d, Teacher=="RF"),
                 prior = prior("uniform(0,1)", class = "Intercept"),
                 family = binomial,
                 sample_prior = T)
# Plotting the predictions of the model (prior + likelihood) against the actual data
pp_check(FlatModel, nsamples = 100)

# plotting the posteriors and the sampling process
plot(FlatModel)


PositiveModel_priorCheck <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = "only")
pp_check(PositiveModel_priorCheck, nsamples = 100)

PositiveModel <- brm(Correct|trials(Questions) ~ 1,
                     data = subset(d, Teacher=="RF"),
                     prior = prior("normal(0.8,0.2)", 
                                   class = "Intercept"),
                     family=binomial,
                     sample_prior = T)
pp_check(PositiveModel, nsamples = 100)
plot(PositiveModel)

SkepticalModel_priorCheck <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior=prior("normal(0.5,0.01)", class = "Intercept"),
                      family=binomial,
                      sample_prior = "only")
pp_check(SkepticalModel_priorCheck, nsamples = 100)

SkepticalModel <- brm(Correct|trials(Questions) ~ 1, 
                      data = subset(d, Teacher=="RF"),
                      prior = prior("normal(0.5,0.01)", class = "Intercept"),
                      family = binomial,
                      sample_prior = T)
pp_check(SkepticalModel, nsamples = 100)
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models
2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.
```{r}
# New data
new_data <- data.frame(
  correct=c(9,8,148,34),
  questions=c(10,12,172,65),
  teacher=c("RF","KT","JS","MW"))

# Making empty dataframe
new_teacher_info <- data.frame(teacher = factor(), MAP = numeric(), chance = numeric(), prior = numeric(), likelihood = numeric(), posterior = numeric())
# Run loop to extract MAP and 'chance above chance' for each teacher
for(i in 1:nrow(new_data)) {
    
    correct <- new_data[i,1]
    questions <- new_data[i,2]
    teacher <- new_data[i,3]
    # Define grid and prior
    prob_grid <- seq(from = 0, to = 1, length.out = 10000)
    prior_sub <- teacher_info %>% filter(teacher == new_data[i,3])
    prior <- prior_sub$posterior
    
    
    # Use my sexy little function
    info <- calc_teacher(teacher, correct, questions, prior, prob_grid) %>% as.data.frame()
    
    names(info)[1] <- "teacher"
    names(info)[2] <- "MAP"
    names(info)[3] <- "chance"
    names(info)[4] <- "prior"
    names(info)[5] <- "likelihood"
    names(info)[6] <- "posterior"
    
    # Combine with premade empty dataframe
    if (nrow(new_teacher_info) == 0) {
      new_teacher_info <- info}
      else {
        new_teacher_info <- rbind(new_teacher_info, info)}
    
    }
#Plot of difference in posterior and prior
plot_teacher <- function(prob_grid, posterior, prior, likelihood, teacher){
  d <- data.frame(grid = prob_grid, posterior = posterior, prior = prior, likelihood = likelihood)
  print(ggplot(d, aes(grid,posterior)) +
          geom_point() +geom_line()+theme_classic() +
          geom_line(aes(grid, prior),color= 'red')+
          geom_line(aes(grid, abs(posterior - prior)), color = 'green')+
          ggtitle(teacher) +
          xlab("Knowledge of CogSci")+ ylab("posterior probability"))
}
for (i in 1:nrow(new_data)){
  #Define input
  prob_grid <- seq(from = 0, to = 1, length.out = 10000)
  plot_sub <- new_teacher_info %>% filter(teacher == new_data[i,3])
  posterior <- plot_sub$posterior
  prior <- plot_sub$prior
  likelihood <- plot_sub$likelihood
  teacher <- new_data[i,3]
  
  #use function
  plot_teacher(prob_grid,posterior,prior,likelihood,teacher)
}

# Difference in MAP
MAP_diff <-new_teacher_info %>% group_by(teacher) %>% 
  summarise(
    MAP_posterior = which.max(posterior),
    MAP_prior = which.max(prior),
    MAP_diff = abs(MAP_posterior - MAP_prior)
  )

formattable(MAP_diff, align = c("l", rep("r", NCOL(MAP_diff) - 1)))

# make a posterior predictive plot - what would my model predict
# how many answer you josh answer correct if here were given 178
# simulate how many correct answer josh would
for (i in 1:nrow(new_data)){
  pos_pred <- rbinom(1e4,  # how many times we simluate this
                  size = [i,2], # how many question he was asked
                  prob = [i,3] == teacher_info$teacher  # samples of teacher's knowledge knowledge 
                  )
  print(pos_pred)
}
hist(pos_pred)
hist(pos_pred-148)
pos_pred <- rbinom(1e4,  # how many times we simluate this
                  size = new_data[1,2], # how many question he was asked
                  prob = new_data[1,3] == teacher_info$teacher  # samples of teacher's knowledge knowledge 
                  )
#for every teacher


 #loop that takes the sum from all teachers
for (i in 1:nrow(new_data)){
sum <- sum(pos_pred >= new_data[i,1]-new_data[i,2]*0.03 & pos_pred <= new_data[i,1]+new_data[i,2]*0.03)/100000 
print(sum)}


prior - posterior # density
# this gives distribution of error of your previous belief compared to your current
# there is a 4.3 % chance that he would 148 correct
  # is this the correct? Yes
  # is this an interesting question? somewhat 
    # hint* not really - what whould more interesting question be?

```

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

